---
title: "Dictionary"
output: 
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

# Functions

## General

1. help(topic) 
Opens the topic in the R-help tab; describes the function and arguments used.
```{r eval=FALSE}
help(mean)
```

2. ?topic
Opens the topic in the R-help tab; describes the function and arguments used.
```{r eval=FALSE}
?mean
```

3. ls()
Lists the objects in the current R environment.
```{r eval=FALSE}
ls()
```

4. dir()
Lists the names of each file in the R project.
```{r eval=FALSE}
dir()
```

5. list.files()
Lists the names of each file in the R project.
```{r eval=FALSE}
list.files()
```

## Input and output

1. save()
Saves an R object to a specified file.
```{r eval=FALSE}
x <- 1
save("x", file="newfile.txt")
```

2. load()
Loads saved data.
```{r eval=FALSE}
load(file="newfile.txt")
```

3. data()
Loads specified data.
```{r eval=FALSE}
data("UCBAdmissions")
```

4. library()
Loads a library/package into R.
```{r eval=FALSE}
library(psych)
```

5. read.table()
Reads a table and produces a data frame.
```{r eval=FALSE}
read.table(file = "newfile.txt")
```

6. read.csv()
Reads a table and produces a data frame.
```{r eval=FALSE}
read.csv(file = "newfile.csv")
```

7. scan()
Read data into a vector from the console or file.
```{r eval=FALSE}
scan(file = "newfile.txt")
```

8. print()
Prints the contents.
```{r}
print("hello")
```

9. cat()
Concatenates and prints the contents.
```{r eval=FALSE}
cat("hello","there")
```

10. write.table()
Converts contents to a data frame and prints.
```{r eval=FALSE}
write.table(x, file = "newfile.txt")
```

## Data Creation

1. c()
Concatonates (combines) data into a vector.
```{r}
c(1,2,3)
```

2. from:to (where from and to are replaced with numbers, e.g. 1:10)
Lists all numbers in between the given numbers.
```{r}
1:10
```

3. seq()
Lists numbers from the first argument to the second argument, adding the third argument to the first argument until the third argument is reached.
```{r}
seq(2,10,3)
```

4. rep()
Replicates the values in the first argument for the amount in the second argument.
```{r}
rep(5, 10)
```

4. x[-(1:n)] elements from n+1 to the end
Returns values of x besides the values in indices 1 to n.
```{r}
x <- 1:10
x[-(1:5)]
```

5. x[c(1,4,2)] specific elements
Returns the values of x in the indices of 1, 4, and 2.
```{r}
x <- 1:10
x[c(1,4,2)]
```

6. <span style="color:red">x["name"] elements named "name"</span>


7. x[x>3] all elements greater than 3
Returns the values of x that are greater than 3.
```{r}
x <- 1:10
x[x>3]
```

8. x[x > 3 & x < 5] all elements between 3 and 5
Returns the va;ues of x that are greater than 3 and less than 5.
```{r}
x <- 1:10
x[x>3 & x<5]
```

9. x[x %in% c("a","and","the")] all elements in given set
Reutrns the values of x that match "a", "and", and "the".
```{r}
x <- c("This","is","the","example","and","the","text","that","will","be","analyzed")
x[x %in% c("a","and","the")]
```

## Indexing lists
1. x[n] list with elements n
Returns the nth object in the list, in list format.
```{r}
x <- list("a" <- c(1,10), "b" <- c(2,20), "c" <- c(3,30))
x[2]
```

2. x[[n]] nt h element of the list
Returns the nth object in the list.
```{r}
x <- list("a" <- c(1,10), "b" <- c(2,20), "c" <- c(3,30))
x[[2]]
```

3. <span style="color:red">x[["name"]] element of the list named "name"</span>


4. <span style="color:red">x$name id.</span>


## Indexing matrices
1. x[i,j] element at row i, column j
Returns the value of x at row i and column j.
```{r}
x <- matrix(c(1,2,3,11,12,13), nrow = 2, ncol = 3, byrow = TRUE)
x[2,2]
```

2. x[i,] row i
Returns all values of x in row i.
```{r}
x <- matrix(c(1,2,3,11,12,13), nrow = 2, ncol = 3, byrow = TRUE)
x[2,]
```

3. x[,j] column j
Returns all values of x in column j.
```{r}
x <- matrix(c(1,2,3,11,12,13), nrow = 2, ncol = 3, byrow = TRUE)
x[,2]
```

4. x[,c(1,3)] columns 1 and 3
Return all values of x in columns 1 and 3 (regardless of rows).
```{r}
x <- matrix(c(1,2,3,11,12,13), nrow = 2, ncol = 3, byrow = TRUE)
x[,c(1,3)]
```

5. <span style="color:red">x["name",] row named "name"</span>


## Indexing data frames (matrix indexing plus the following) 
1. <span style="color:red>x[["name"]] column named "name"</span>

2. x$nameid.
Returns column of x labeled "nameid".
```{r}
a <- c(1,10,100)
b <- c(2,20,200)
c <- c(3,30,300)
x <- data.frame(a,b,c)
x$a
```

## Variable conversion

1. as.data.frame(x)
Transforms object into a data frame.
```{r}
x <- as.data.frame(matrix(c(1,2,3,11,12,13), nrow = 2, ncol = 3, byrow = TRUE))
x
```

2. as.numeric(x)
Transforms object into numeric class (if possible).
```{r}
x <- "1"
x <- as.numeric(x)
class(x)
```

3. as.logical(x)
Transforms object into boolean class (if possible).
```{r}
x <- 1
x <- as.logical(x)
class(x)
```

4. as.character(x)
Transforms object into character class.
```{r}
x <- 1
x <- as.character(x)
class(x)
```

## Variable information
1. is.na(x)
2. is.null(x)
3. is.data.frame(x)
4. is.numeric(x)
5. is.character(x)
6. length(x)
7. dim(x)
8. dimnames(x)
9. nrow(x)
10. ncol(x)
11. class()
12. attributes()

## Data selection and manipulation
1. which.max()
2. which.min()
3. which()
4. sort()
5. unique()
6. table()
7. sample()

## Math
1. max()
2. min()
3. range()
4. sum()
5. mean()
6. median()
7. var()
8. sd()
9. cor()
10. round()
11. abs()

## Matrices
1. t()
2. diag()
3. rowSums()
4. colSums()
5. rowMeans()
6. colMeans()

## Advanced Data processing
1. apply()
2. aggregate()

## Strings
1. paste()
2. strsplit()
3. tolower()
4. toupper

## Plotting
1. hist()
2. plot()

## Distributions
1. rnorm()
2. runif()

## Programming
1. show that you can define a function
2. show that you can write a for loop
3. show that you can write a while loop
4. show that you can write an if else statement
5. Explain what return() does inside a function, show you can use it
6. Explain what break() does, show you can use it

## Installing some packages

Use the packages tab in R-studio to install these packages. You will need to be connected to the internet when you do this. If you are installing on your laptop, or on R-studio Cloud, then these packages will not need to installed again. 

1. ggplot2
2. dplyr
3. shiny
4. data.table
5. reshape2
6. stringr

