---
title: "Jeff's Journal"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 1
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

## Notes to myself:
Option+Command+i to insert R code in HTML  
Control+Enter to run inserted R code  
{r eval=false} to insert R code without running it  

# Easy Problems

## 1. Simple math with numbers
```{r}
1+68
50-3
9/3
142*5
```

## 2. Put numbers into variables, do simple math
```{r}
jawbreakers = 10
cost = 25
jawbreakers * cost
cost/jawbreakers
```

# 3a. Place the numbers 1 to 100 separately into a variable using for loop
```{r}
ed <- length(100)
for (i in 1:100) {
  ed[i] <- i
}
print(ed)
```

## 3b. Place the numbers 1 to 100 separately into a variable using seq
```{r}
edd <- seq(1,100,1)
print(edd)
```

## 4. Find the sum of all the integer numbers from 1 to 100
```{r}
eddie <- sum(seq(1,100,1))
print(eddie)
```

## 5. Write a function to find the sum of all integers between any two values
```{r}
int_sum <- function(x, y) {
  temp = 0
  sum = 0
  if (x > y){
    temp = x
    x = y
    y = temp
  }
  for (i in x:y){
    sum = sum + i
  }
  print(sum)
}
```

