---
title: "Jeff's Journal"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 2
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

# Easy Problems

## 1. Simple math with numbers
```{r}
1+68
50-3
9/3
142*5
```

## 2. Put numbers into variables, do simple math
```{r}
jawbreakers = 10
cost = 25
jawbreakers * cost
cost/jawbreakers
```

## 3a. Place the numbers 1 to 100 separately into a variable using for loop
```{r}
ed <- length(100)
for (i in 1:100) {
  ed[i] <- i
}
print(ed)
```

## 3b. Place the numbers 1 to 100 separately into a variable using seq
```{r}
edd <- seq(1,100,1)
print(edd)
```

## 4. Find the sum of all the integer numbers from 1 to 100
```{r}
eddie <- sum(seq(1,100,1))
print(eddie)
```

## 5. Write a function to find the sum of all integers between any two values
```{r}
int_sum <- function(x, y) {
  temp = 0
  jeff_sum = 0
  #Switch x and y if (x > y) -- enables for loop
  if (x > y){
    temp = x
    x = y
    y = temp
  }
  for (i in x:y){
    jeff_sum = jeff_sum + i
  }
  return(jeff_sum)
}
```

## 6. List all of the odd numbers from 1 to 100
```{r}
odd_sum <- sum(seq(1,100,2))
print(odd_sum)
```

## 7. List all of the prime numbers from 1 to 1000
```{r}
prime_numbers <- c()
for (i in 3:1000) {
  prime = TRUE
  for (j in 2:(i-1)) {
    if (i %% j == 0) {
      prime = FALSE
    }
  }
  if (prime == TRUE) {
    prime_numbers <- c(prime_numbers,i)
  }
}
print(prime_numbers)
```

## 8. Generate 100 random numbers
```{r}
sample(1:100, 100, replace=T)
```

## 9. Generate 100 random numbers within a specific range
```{r}
sample(1:100, 100, replace=T)
```

## 10a. Write your own functions to give descriptive statistics for a vector variable: Mean
```{r}
x <- c(1,2,3,4,5,6,7,8,9)
jeff_mean <- function(x) {
  func_mean <- 0
  for (i in 1:length(x)) {
    func_mean <- func_mean + (x[i]/length(x))
  }
  return(func_mean)
}
jeff_mean(x)
```

## 10b. Write your own functions to give descriptive statistics for a vector variable: Mode
```{r}
x <- c(1,1,1,1,2,2,3,3,3,4,4,4,4)
jeff_mode <- function(x) {
  y <- length(length(x))
  mode_list <- c()
  x_copy <- x
  for (i in 1:length(x)) {
    y[i] <- length(which(x %in% x[i]))
  }
  max_count_pos <- match(max(y),y)
  max_count <- max(y)
  #Check for multiple modes
  for (i in 1:length(x)) {
    if (y[i] == max_count) {
      if ((x[i] %in% x_copy[-1]) == FALSE) {
        mode_list <- c(mode_list,x[i])
      }
    }
    x_copy <- x_copy[-1]
  }
  return(mode_list)
}
jeff_mode(x)
```

## 10c. Write your own functions to give descriptive statistics for a vector variable: Median
```{r}
x <- c(1,2,3,4,5,6,7,8,9)
jeff_median <- function(x) {
  while (length(x) > 2) {
    loc <- match(c(max(x),min(x)),x)
    x <- x[-loc]
    }
    if (length(x) == 2) {
    x <- (x[1]+x[2])/2
    }
  return(x)
}
jeff_median(x)
```

## 10d. Write your own functions to give descriptive statistics for a vector variable: Range
```{r}
x <- c(1,2,3,4,5,6,7,8,9)
jeff_range <- function(x) {
  func_range <- max(x)-min(x)
  return(func_range)
}
jeff_range(x)
```

## 10d. Write your own functions to give descriptive statistics for a vector variable: Standard Deviation
```{r}
x <- c(1,2,3,4,5,6,7,8,9)
jeff_sd <- function(x) {
  func_mean <- mean(x)
  var <- 0
  for (i in 1:length(x)) {
    var <- var + (func_mean - x[i])^2
    var <- var/length(x)
    stdev <- sqrt(var)
  }
  return(stdev)
}
jeff_sd(x)
```

## 11. Count the number of characters in a string variable
```{r}
hi <- "hello there"
nchar(hi)
```

## 12. Count the number of words in a string variable
```{r}
name <- "my name jeff"
word_count <- length(strsplit(name,' ')[[1]])
print(word_count)
```

## 13. Count the number of sentences in a string variable
```{r}
alice <- "If I had a world of my own, everything would be nonsense. Nothing would be what it is, because everything would be what it isn't. And contrary wise, what is, it wouldn't be. And what it wouldn't be, it would."
sentence_count <- length(strsplit(alice,"\\.")[[1]])
print(sentence_count)
strsplit(alice,'\\.')[[1]]
```

## 14. Count the number of times a specific character occurs in a string variable
```{r}
alice <- "Curiouser and curiouser!"
find <- "u"
count <- length(which(strsplit(alice, "")[[1]]==find))
print(count)
```

