---
title: "Jeff's Journal"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 2
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

# Easy Problems

## 1. Simple math with numbers
```{r}
1+68
50-3
9/3
142*5
```

## 2. Put numbers into variables, do simple math
```{r}
jawbreakers = 10
cost = 25
jawbreakers * cost
cost/jawbreakers
```

## 3a. Place the numbers 1 to 100 separately into a variable using for loop
```{r}
ed <- length(100)
for (i in 1:100) {
  ed[i] <- i
}
print(ed)
```

## 3b. Place the numbers 1 to 100 separately into a variable using seq
```{r}
edd <- seq(1,100,1)
print(edd)
```

## 4. Find the sum of all the integer numbers from 1 to 100
```{r}
eddie <- sum(seq(1,100,1))
print(eddie)
```

## 5. Write a function to find the sum of all integers between any two values
```{r}
int_sum <- function(x, y) {
  jeff_sum = 0
  #Switch x and y if (x > y) -- enables for loop
  if (x > y){
    temp = x
    x = y
    y = temp
  }
  for (i in x:y){
    jeff_sum = jeff_sum + i
  }
  return(jeff_sum)
}
int_sum(5,1)
```

## 6. List all of the odd numbers from 1 to 100
```{r}
odd_nums <- seq(1,100,2)
print(odd_nums)
```

## 7. List all of the prime numbers from 1 to 1000
```{r}
prime_numbers <- c()
for (i in 3:1000) {
  prime = TRUE
  for (j in 2:(i-1)) {
    if (i %% j == 0) {
      prime = FALSE
    }
  }
  if (prime == TRUE) {
    prime_numbers <- c(prime_numbers,i)
  }
}
print(prime_numbers)
```

## 8. Generate 100 random numbers
```{r}
sample(1:100, 100, replace=T)
```

## 9. Generate 100 random numbers within a specific range
```{r}
sample(1:100, 100, replace=T)
```

## 10a. Write your own functions to give descriptive statistics for a vector variable: Mean
```{r}
x <- c(1,2,3,4,5,6,7,8,9)
jeff_mean <- function(x) {
  func_mean <- 0
  for (i in 1:length(x)) {
    func_mean <- func_mean + (x[i]/length(x))
  }
  return(func_mean)
}
jeff_mean(x)
```

## 10b. Write your own functions to give descriptive statistics for a vector variable: Mode
```{r}
x <- c(1,1,1,1,2,2,3,3,3,4,4,4,4)
jeff_mode <- function(x) {
  y <- c()
  mode_list <- c()
  x_copy <- x
  for (i in 1:length(x)) {
    y[i] <- length(which(x %in% x[i]))
  }
  max_count_pos <- match(max(y),y)
  max_count <- max(y)
  #Check for multiple modes
  for (i in 1:length(x)) {
    if (y[i] == max_count) {
      if ((x[i] %in% x_copy[-1]) == FALSE) {
        mode_list <- c(mode_list,x[i])
      }
    }
    x_copy <- x_copy[-1]
  }
  return(mode_list)
}
jeff_mode(x)
```

## 10c. Write your own functions to give descriptive statistics for a vector variable: Median
```{r}
x <- c(1,2,3,4,5,6,7,8,9)
jeff_median <- function(x) {
  while (length(x) > 2) {
    loc <- match(c(max(x),min(x)),x)
    x <- x[-loc]
    }
    if (length(x) == 2) {
    x <- (x[1]+x[2])/2
    }
  return(x)
}
jeff_median(x)
```

## 10d. Write your own functions to give descriptive statistics for a vector variable: Range
```{r}
x <- c(1,2,3,4,5,6,7,8,9)
jeff_range <- function(x) {
  func_range <- max(x)-min(x)
  return(func_range)
}
jeff_range(x)
```

## 10d. Write your own functions to give descriptive statistics for a vector variable: Standard Deviation
```{r}
x <- c(1,2,3,4,5,6,7,8,9)
jeff_sd <- function(x) {
  func_mean <- mean(x)
  var <- 0
  for (i in 1:length(x)) {
    var <- var + (func_mean - x[i])^2
    var <- var/length(x)
    stdev <- sqrt(var)
  }
  return(stdev)
}
jeff_sd(x)
```

## 11. Count the number of characters in a string variable
```{r}
hi <- "hello there"
nchar(hi)
```

## 12. Count the number of words in a string variable
```{r}
name <- "my name jeff"
word_count <- length(strsplit(name,' ')[[1]])
print(word_count)
```

## 13. Count the number of sentences in a string variable
```{r}
alice <- "If I had a world of my own, everything would be nonsense. Nothing would be what it is, because everything would be what it isn't. And contrary wise, what is, it wouldn't be. And what it wouldn't be, it would."
sentence_count <- length(strsplit(alice,"\\.")[[1]])
print(sentence_count)
strsplit(alice,'\\.')[[1]]
```

## 14. Count the number of times a specific character occurs in a string variable
```{r}
alice <- "Curiouser and curiouser!"
find <- "u"
count <- length(which(strsplit(alice, "")[[1]]==find))
print(count)
```

## 15. Do a logical test to see if one word is found within the text of another string variable
```{r}
alice <- "If I had a world of my own, everything would be nonsense. Nothing would be what it is, because everything would be what it isn't. And contrary wise, what is, it wouldn't be. And what it wouldn't be, it would."
find <- "everything"
exist <- grep(find, alice)
print(exist)
```

## 16. Put the current computer time in milliseconds into a variable
```{r}
current_time <- as.numeric(Sys.time())
print(current_time)
```

## 17. Measure how long a piece of code takes to run by measuring the time before the code is run, and after the code is run, and taking the difference to find the total time
```{r}
start_time <- as.numeric(Sys.time())
## Start example code
ed <- length(100)
for (i in 1:100) {
  ed[i] <- i
}
## End example code
end_time <- as.numeric(Sys.time())
run_time <- end_time - start_time
print(run_time)
```

## 18. Read a .txt file or .csv file into a variable
```{r eval=FALSE}
txt_data <- scan(file = "insert_file_here.txt")
csv_data <- csv.read(file = "insert_file_here.csv")
```


## 19. Output the contents of a variable to a .txt file
```{r}
jeff_file <- file("output.txt")
writeLines(c("Hello","World"), jeff_file)
close(jeff_file)
```

## 20. Create a variable that stores a 20x20 matrix of random numbers
```{r}
data <- replicate(20, runif(20, min = 0, max = 100))
print(data)
```


## 21. Output any matrix to a txt file using commas or tabs to separate column values, and new lines to separate row values
```{r}
data <- replicate(20, runif(20, min = 0, max = 100))
write.csv(data, file = "data_comma.csv", sep = ",")
write.csv(data, file = "data_tab.csv", sep = "\t")
```


# Hard Problems

## 1. The FizzBuzz Problem
List the numbers from 1 to 100 with the following constraints. If the number can be divided by three evenly, then print Fizz instead of the number. If the number can be divided by five evenly, then print Buzz instead of the number. Finally, if the number can be divided by three and five evenly, then print FizzBuzz instead of the number.
```{r}
total <- 1
for (i in 2:100) {
  if (i %% 3 == 0) {
    if (i %% 5 == 0) {
      total <- paste(total,"FizzBuzz", sep = ", ")
    }
    else {
      total <- paste(total,"Fizz", sep = ", ")
    }
  }
  else if (i %% 5 == 0) {
    total <- paste(total, "Buzz", sep = ", ")
  }
  else {
    total <- paste(total, i, sep = ", ")
  }
}
print(total)
```

## 2. Frequeny Counts
Take text as input, and be able to produce a table that shows the counts for each character in the text. This problem is related to the earlier easy problem asking you to count the number of times that a single letter appears in a text. The slightly harder problem is the more general version: count the frequencies of all unique characters in a text.
```{r}
alice <- "curiouser and curiouser!"
table(unlist(strsplit(alice, split = "")))
```

Without using table:
```{r}
alice <- "curiouser and curiouser!"
letters <- unique(unlist(strsplit(alice, split="")))
count <- c()
for (i in 1:length(letters)) {
  count[i] <- length(which(strsplit(alice, "")[[1]]==letters[i]))
}
alice_data <- data.frame(count,letters)
print(alice_data)
```


## 3. Test the Random Number Generator
Test the random number generator for a flat distribution. Generate a few million random numbers between 0 and 100. Count the number of 0s, 1s, 2s, 3s, etc. all the way up to 100. Look at the counts for each of the numbers and determine if they are relatively equal. For example, you could plot the counts in Excel to make a histogram. If all of the bars are close to being flat, then each number had an equal chance of being selected, and the random number generator is working without bias.
```{r}
numbers <- round(runif(3000000, min = 0, max = 100))
count <- c()
for (i in 1:100) {
  count[i] <- length(which(numbers %in% i))
}
print(count)
plot(count)
```


## 4. Create a Multiplication Table
Generate a matrix for a multiplication table. For example, the labels for the columns could be the numbers 1 to 10, and the labels for the rows could be the numbers 1 to 10. The contents of each of the cells in the matrix should be correct answer for multiplying the column value by the row value.
```{r}
x <- seq(1, 10, 1)
data <- matrix(x, nrow = 10, ncol = 10)
for (i in 1:10) {
  x <- seq(i, i*10, i)
  data[,i] <- x
}
#data <- matrix(x, nrow = 10, ncol = 1)
print(data)
```

## 5. Encrypt and Decrypt the Alphabet
Turn any normal english text into an encrypted version of the text, and be able to turn any decrypted text back into normal english text. A simple encryption would be to scramble the alphabet such that each letter corresponds to a new randomly chosen (but unique) letter.
```{r}
# Initialize text and randomized keycode
alice <- "curiouser and curiouser"
split_alice <- unlist(strsplit(alice, split=""))
alphabet <- c("a","b","c","d","e","f","g","h","i","j","k","l","m","n","o","p","q","r","s","t","u","v","w","x","y","z")
coded_alice <- c()
keycode <- data.frame(c(1:26),alphabet,sample(1:26, 26, replace = FALSE))
print(alice)
# Translate each character into the coded equivalent
for (i in 1:length(split_alice)) {
  for (j in 1:26) {
    if (split_alice[i] == keycode[j,2]) {
      coded_alice <- paste(coded_alice, as.character(keycode[keycode[j,3],2]), sep="")
    }
    else if (split_alice[i] == " "){
      coded_alice <- paste(coded_alice, "")
      break
    }
  }
}
print(coded_alice)
# Decode text
coded_alice <- unlist(strsplit(coded_alice, split=""))
decoded_alice <- c()
for (i in 1:length(coded_alice)) {
  for (j in 1:26) {
    if (coded_alice[i] == keycode[j,2]) {
      for (k in 1:26) {
        if (keycode[k,3] == keycode[j,1]) {
          decoded_alice <- paste(decoded_alice, as.character(keycode[k,2]), sep="")
        }
      }
    }
    else if (coded_alice[i] == " "){
      decoded_alice <- paste(decoded_alice, "")
      break
    }
  }
}
print(decoded_alice)
```

## 6. Snakes and Ladders
Your task here is to write an algorithm that can simulate playing the above depicted Snakes and Ladders board. You should assume that each roll of the dice produces a random number between 1 and 6. After you are able to simulate one played game, you will then write a loop to simulate 1000 games, and estimate the average number of dice rolls needed to successfully complete the game.
```{r}
move <- 0
count <- replicate(1000, 0)
for (i in 1:1000) {
  spot <- 1
  while (spot <= 25) {
  move <- sample(1:6, 1)
  spot <- spot + move
  if (spot == 3) {
    spot <- 11
  }
  else if (spot == 6) {
    spot <- 17
  }
  else if (spot == 9) {
    spot <- 18
  }
  else if (spot == 10) {
    spot <- 12
  }
  else if (spot == 14) {
    spot <- 4
  }
  else if (spot == 19) {
    spot <- 8
  }
  else if (spot == 22) {
    spot <- 20
  }
  else if (spot == 24) {
    spot <- 16
  }
  count[i] = count[i] + 1
  }
}
avg_moves <- mean(count)
print(avg_moves)
```

With shorter code:
```{r}
move <- 0
count <- replicate(1000, 0)
game_grid <- data.frame(c(3,6,9,10,14,19,22,24),c(11,17,18,12,4,8,20,16))
for (i in 1:1000) {
  spot <- 1
  while (spot <= 25) {
  move <- sample(1:6, 1)
  spot <- spot + move
  for (i in 1:8) {
    if (spot == game_grid[i,1]) {
      spot <- game_grid[i,2]
    }
  }
  count[i] = count[i] + 1
  }
}
avg_moves <- mean(count)
print(avg_moves)
```

## 7. Dice Rolling Simulations
Assume that a pair of dice are rolled. Using monte carlo-simulation, compute the probabilities of rolling a 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, and 12, respectively.
```{r}
prob <- data.frame(c(2:12),replicate(11,0))
for (i in 1:10000) {
  roll <- sum(sample(1:6, 2, replace = TRUE))
  for (j in 2:12) {
    if (roll == j) {
      prob[j-1,2] = prob[j-1,2] + 1
    }
  }
}
prob[,2] = prob[,2]/10000
print(prob)
```

## 8. Monte Hall Problem
The monte-hall problem is as follows. A contestant in a game show is presented with three closed doors. They are told that a prize is behind one door, and two goats are behind the other two doors. They are asked to choose which door contains the prize. After making their choice the game show host opens one of the remaining two doors (not chosen by the contestant), and reveals a goat. There are now two door remaining. The contestant is asked if they would like to switch their choice to the other door, or keep their initial choice. The correct answer is that the participant should switch their initial choice, and choose the other door. This will increase their odds of winning. Demonstrate by monte-carlo simulation that the odds of winning is higher if the participant switches than if the participants keeps their original choice.
```{r}
choice <- c("Stay","Switch")
probability_of_winning <- c(0,0)
prob <- data.frame(choice,probability_of_winning)
for (i in 1:10000) {
  stay_prob <- sample(1:3, 1)
  if (stay_prob == 1) {
    prob[1,2] <- prob[1,2] + 1
  }
  switch_prob <- sample(1:2, 1)
  if (switch_prob == 1) {
    prob[2,2] <- prob[2,2] + 1
  }
}
prob[,2] <- prob[,2]/10000
print(prob)
```

## 9. 100 Doors Problem
Problem: You have 100 doors in a row that are all initially closed. You make 100 passes by the doors. The first time through, you visit every door and toggle the door (if the door is closed, you open it; if it is open, you close it). The second time you only visit every 2nd door (door 2, 4, 6, etc.). The third time, every 3rd door (door 3, 6, 9, etc.), etc, until you only visit the 100th door.
```{r}
# Creates 100 doors, 0 indicating that a door is closed
doors <- replicate(100, 0)
for (i in 1:100) {
  j = i
  k = i
  # For each door, if/else open closed doors or closes open doors
  while (j <= 100) {
    if (doors[j] == 0) {
      doors[j] <- 1
    }
    else {
      doors[j] <- 0
    }
    # Check only the doors that are multiples of the first door checked
    j = j + k
  }
}
print(doors)
```

## 10. 99 Bottles of Beer Problem
In this puzzle, write code to print out the entire “99 bottles of beer on the wall”" song.
```{r}
i <- 99
lyrics_1 <- "bottles of beer on the wall,"
lyrics_2 <- "bottles of beer. Take one down, pass it around,"
lyrics_3 <- "bottles of beer on the wall!"
while (i > 0) {
  each_lyric <- paste(i,lyrics_1,i,lyrics_2,i-1,lyrics_3)
  print(each_lyric)
  i = i -1
}
```

## 11. Random Tic-Tac-Toe
Imagine that two players make completely random choices when playing tic-tac-toe. Each game will either end in a draw or one of the two players will win. Create a monte-carlo simulation of this “random” version of tic-tac-toe. Out 10,000 simulations, what proportion of the time is the game won versus drawn?

Raw code:
```{r}
win_count <- 0
winning_pos <- data.frame(c(1,2,3),c(1,4,7),c(1,5,9),c(4,5,6),c(7,8,9),c(2,5,8),c(3,6,9),c(3,5,7))
end_loop <-FALSE
for (i in 1:10000) {
  count_p1 <- 0
  count_p2 <- 0
  board <- sample(1:9, 9, replace=FALSE)
  board_p1 <- board[c(1,3,5,7,9)]
  board_p2 <- c(board[c(2,4,6,8)],0)
  for (j in 1:8) {
    for (k in 1:3) {
      for (l in 1:5) {
        if (board_p1[l] == winning_pos[k,j]) {
          count_p1 <- count_p1 + 1
        }
        if (board_p2[l] == winning_pos[k,j]) {
          count_p2 <- count_p2 + 1
        }
      }
      if (count_p1 == 3) {
        win_count <- win_count + 1
        end_loop <- TRUE
        break
      }
      if (count_p2 == 3) {
        win_count <- win_count + 1
        end_loop <- TRUE
        break
      }
    }
    count_p1 <- 0
    count_p2 <- 0
    if (end_loop == TRUE) {
      end_loop <- FALSE
      break
    }
  }
}
win_prob <- win_count/10000
print(win_prob)
```

Annotated for ease of readibility:
```{r eval=FALSE}
# Count the number of times the game is won
win_count <- 0
# Create a data frame of each of the possible ways one can win
winning_pos <- data.frame(c(1,2,3),c(1,4,7),c(1,5,9),c(4,5,6),c(7,8,9),c(2,5,8),c(3,6,9),c(3,5,7))
# Initialize a variable that signals breaks from multiple nested loops
end_loop <-FALSE
for (i in 1:10000) {
  # Initialize a count of completed necessary moves to win for each player; 
  # reset to 0 for each simulation
  count_p1 <- 0
  count_p2 <- 0
  # Creates a randomly filled board; no spot can be filled twice
  board <- sample(1:9, 9, replace=FALSE)
  # Assigns every other move to player 1 and player 2;
  # player 2 concatenation with 0 enables vector to be used in for(l in 1:5) loop
  board_p1 <- board[c(1,3,5,7,9)]
  board_p2 <- c(board[c(2,4,6,8)],0)
  # Check each possible way to win
  for (j in 1:8) {
    # Check if each of the three required moves is filled
    for (k in 1:3) {
      # For each of player 1's moves, check if it matches one of the 3 required moves
      for (l in 1:5) {
        # If the player 1 move matches a required move, add to the player count
        if (board_p1[l] == winning_pos[k,j]) {
          count_p1 <- count_p1 + 1
        }
        # If the player 1 move matches a required move, add to the player count
        if (board_p2[l] == winning_pos[k,j]) {
          count_p2 <- count_p2 + 1
        }
      }
      # If all 3 required moves to win are completed, add to win count,
      # assign end_loop to TRUE to signal to break out of the for(j in 1:8) loop,
      # and break out of the for(k in 1:3) loop
      if (count_p1 == 3) {
        win_count <- win_count + 1
        end_loop <- TRUE
        break
      }
      # The above break ensures that if the board is randomly filled so that both 
      # players win, the if statement below will not run, and win count will not 
      # be increased again
      if (count_p2 == 3) {
        win_count <- win_count + 1
        end_loop <- TRUE
        break
      }
    }
    # Resets completed necessary moves to 0; this ensures that 2 required moves from
    # the first possible way to win are not carried over to the second possible way to win
    count_p1 <- 0
    count_p2 <- 0
    # Breaks the for(j in 1:8) loop; once a winning combination is found, the rest will not
    # be checked. This ensures that if the game is won in 2 different ways, it will only 
    # be counted as 1 win
    if (end_loop == TRUE) {
      end_loop <- FALSE
      break
    }
  }
}
# Divides the win count by 10,000 to create a probability that the game is won
win_prob <- win_count/10000
print(win_prob)
```