---
title: "Jeff's Journal"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 2
    #code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```

# Easy Problems

## 1. Simple math with numbers
```{r}
1+68
50-3
9/3
142*5
```

## 2. Put numbers into variables, do simple math
```{r}
jawbreakers = 10
cost = 25
jawbreakers * cost
cost/jawbreakers
```

## 3a. Place the numbers 1 to 100 separately into a variable using for loop
```{r}
ed <- length(100)
for (i in 1:100) {
  ed[i] <- i
}
print(ed)
```

## 3b. Place the numbers 1 to 100 separately into a variable using seq
```{r}
edd <- seq(1,100,1)
print(edd)
```

## 4. Find the sum of all the integer numbers from 1 to 100
```{r}
eddie <- sum(seq(1,100,1))
print(eddie)
```

## 5. Write a function to find the sum of all integers between any two values
```{r}
int_sum <- function(x, y) {
  temp = 0
  jeff_sum = 0
  #Switch x and y if (x > y) -- enables for loop
  if (x > y){
    temp = x
    x = y
    y = temp
  }
  for (i in x:y){
    jeff_sum = jeff_sum + i
  }
  return(jeff_sum)
  print(jeff_sum)
}
```

## 6. List all of the odd numbers from 1 to 100
```{r}
odd_sum <- sum(seq(1,100,2))
print(odd_sum)
```

## 7. List all of the prime numbers from 1 to 1000
```{r}
prime_numbers <- c()
for (i in 3:1000) {
  prime = TRUE
  for (j in 2:(i-1)) {
    if (i %% j == 0) {
      prime = FALSE
    }
  }
  if (prime == TRUE) {
    prime_numbers <- c(prime_numbers,i)
  }
}
print(prime_numbers)
```


## 8. Generate 100 random numbers
```{r}
sample(1:100, 100, replace=T)
```

## 9. Generate 100 random numbers within a specific range
```{r}
sample(1:100, 100, replace=T)
```

## 10a. Write your own functions to give descriptive statistics for a vector variable: Mean
```{r}
x <- c(1,2,3,4,5,6,7,8,9)
jeff_mean <- function(x) {
  func_mean <- 0
  for (i in 1:(length(x))) {
    func_mean <- func_mean + (x[i]/length(x))
  }
  return(func_mean)
  print(func_mean)
}
jeff_mean(x)
```

## 10b. Write your own functions to give descriptive statistics for a vector variable: Mode
```{r}
x <- c(1,1,1,1,2,2,3,3,3,4,4,4,4)
jeff_mode <- function(x) {
  y <- length(length(x))
  mode_list <- c()
  x_copy <- x
  for (i in 1:length(x)) {
    y[i] <- length(which(x %in% x[i]))
  }
  max_count_pos <- match(max(y),y)
  max_count <- max(y)
  #Check for multiple modes
  for (i in 1:(length(x))) {
    if (y[i] == max_count) {
      if ((x[i] %in% x_copy[-1]) == FALSE) {
        mode_list <- c(mode_list,x[i])
      }
    }
    x_copy <- x_copy[-1]
  }
  return(mode_list)
  print(mode_list)
}
jeff_mode(x)
```

## 10c. Write your own functions to give descriptive statistics for a vector variable: Median
```{r}
x <- c(1,2,3,4,5,6,7,8,9)
jeff_median <- function(x) {
  while (length(x) > 2) {
    loc <- match(c(max(x),min(x)),x)
    x <- x[-loc]
    }
    if (length(x) == 2) {
    x <- (x[1]+x[2])/2
    }
  print(x)
  return(x)
}
jeff_median(x)
```

