---
title: "Resources"
output: 
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: false
    toc_depth: 2
---

# Additional R resources

## Crump

1. [Matt Crump's new stats book](https://crumplab.github.io/statistics/)

2. [Matt Crump's older programming book](https://crumplab.github.io/programmingforpsych/)

3. [Matt Crump's lab manual](https://crumplab.github.io/statisticsLab/)

## Open source textbooks

1. [Danielle Navarro's free Psych Stats textbook using R](https://compcogscisydney.org/learning-statistics-with-r/)

2. [Daniell Navarro's recently made website for introducing R](http://compcogscisydney.org/psyr/)

3. [Guide for R Markdown](https://bookdown.org/yihui/rmarkdown/)

## Other

1. [Stackoverflow](https://stackoverflow.com) 

2. [Journal Template](https://crumplab.github.io/LabJournalWebsite/Journal.html)

# Notes (for mac)

1. **Insert R code in HTML**: Option+Command+i  
2. **Run inserted R code:** Control+Enter  
3. **Insert R code without running it:** {r eval=false}
4. **Insert linked text:** 
```{r eval=FALSE}
[text](link)
```
5. **Change color of text:** 
```{r eval=FALSE}
<span style="color:red">TEXT HERE</span>
```

